{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-12-26 thesis/landslide.tex
footnote abt no-memory return-value yield-blocking loops

In order to detect when a thread should be unblocked from its yield loop,
Landslide simply leverages DPOR's existing computation of memory conflicts:
whatever condition the blocked thread was waiting for
will show up in the conflicts between it and whichever thread fulfills it.
Hence, every memory conflict detected during DPOR is also checked against any
currently yield-loop-blocked threads, unblocking them in the case of a match
({\tt check\_unblock\_yield\_loop()} in {\tt user\_sync.c}).
If that memory access is not sufficient to let the blocked thread start making progress again,
it will simply trigger the yield-blocking heuristic again.
In theory, this could result in a livelock between two threads,
each in principle blocked in a yield loop,
but where the conditions of the loop happen to conflict with each other,
causing the threads to keep waking each other up;
however,
% hmm... is this true? like if they use mutexes on each other or smth, or use HTM to check a condition
% where the htm failure path uses a stop the world thing...
I have never observed this in practice, as the conditions \revisionminor{checked by} such blocking loops are usually read-only.
Future work could heuristically address this by deprioritizing such threads,
as measured by the number of times they've yield-blocked,
so that a third thread which could actually make progress may run if it exists.

>>>>

In order to detect when a thread should be unblocked from its yield loop,
Landslide simply leverages DPOR's existing computation of memory conflicts:
whatever condition the blocked thread was waiting for
will show up in the conflicts between it and whichever thread fulfills it.%
\footnote{A yield-blocking loop may depend on another thread's execution with no memory conflicts at all
simply by checking the return value to observe whether the yield was successful.
In principle, Landslide should check for such loops whenever the yielded-to thread's runnability changes;
at present, this is not implemented,
and instead leans on the false positive deadlock detector to wake the blocking thread up.}
Hence, every memory conflict detected during DPOR is also checked against any
currently yield-loop-blocked threads, unblocking them in the case of a match
({\tt check\_unblock\_yield\_loop()} in {\tt user\_sync.c}).
If that memory access is not sufficient to let the blocked thread start making progress again,
it will simply trigger the yield-blocking heuristic again.
In theory, this could result in a livelock between two threads,
each in principle blocked in a yield loop,
but where the conditions of the loop happen to conflict with each other,
causing the threads to keep waking each other up;
however,
% hmm... is this true? like if they use mutexes on each other or smth, or use HTM to check a condition
% where the htm failure path uses a stop the world thing...
I have never observed this in practice, as the conditions \revisionminor{checked by} such blocking loops are usually read-only.
Future work could heuristically address this by deprioritizing such threads,
as measured by the number of times they've yield-blocked,
so that a third thread which could actually make progress may run if it exists.
